Numerical simulation and entropy dissipative cure of the carbuncle instability for the shallow water circular hydraulic jump

### 2020-11-19 numerics_paper/version2/numerical_chj.tex
Editorial revisions.

\newcommand{\W}{{\mathcal W}}
\newcommand{\A}{{\mathcal A}}
\newcommand{\bff}{{\bf f}}
\newcommand{\bfr}{{\bf r}}
\newcommand{\bfF}{{\bf F}}
\newcommand{\bfu}{{\bf u}}
\newcommand{\bfv}{{\bf v}}
\newcommand{\bfq}{{\bf q}}
\newcommand{\bfx}{{\bf x}}
\newcommand{\sgn}{{\rm sgn}}
\newcommand{\Rus}{{\rm Rus}}
\newcommand{\Roe}{{\rm Roe}}

@@

\section{The shallow water circular hydraulic jump}
The shallow water equations in two (horizontal) dimensions are
\begin{subequations} \label{eq:sw}
\begin{align}
    h_t + (hu)_x + (hv)_y & = 0 \\
    (hu)_t + \left(hu^2 + \frac{1}{2}gh^2\right)_x + (huv)_y & = 0 \\
    (hv)_t + (huv)_x + \left(hv^2 + \frac{1}{2}gh^2\right)_y & = 0.
\end{align}
\end{subequations}
These can be written in vector form as
\begin{align}
    q_t + f(q)_x + g(q)_y & = 0.
\end{align}

\subsection{Semi-analytical steady solution under rotational symmetry}\label{sec:steady_chj}

By assuming rotational symmetry in \eqref{eq:sw}, one obtains the
system (in one spatial variable)
\begin{subequations} \label{eq:rsw}
\begin{align}
    (rh)_t + (rhu)_r & = 0 \label{mass1} \\
    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r = 0. \label{mom1}
\end{align}
\end{subequations}
Steady-state solutions of \eqref{eq:rsw} satisfy
\begin{subequations}\label{steady}
\begin{align}
    rhu & = C \\
    h'(r) & = \frac{h}{\frac{g}{\beta^2} r^3 h^3 -r} = \frac{h}{r} \cdot \frac{F^2}{1-F^2} \label{eq:dh}
\end{align}
\end{subequations}
for some $C$ independent of $r$.  Here $F=|u|/\sqrt{gh}$ is the Froude number.
We see that two types of steady profiles exist, depending on whether the flow
is subcritical ($|F|<1$) or supercritical ($|F|>1$).  No smooth steady solution can
include both regimes, since the right hand side of \eqref{eq:dh} blows up when $F=1$.

@@

A steady-state, rotationally symmetric solution can be given for an annular region as follows:

@@

Numerical tests suggest that the solution obtained in this way is a stable equilibrium
of the reduced one-dimensional shallow water equations \eqref{eq:rsw}.  This
solution provides a useful starting point for testing the stability of the CHJ
as a solution of the full two-dimensional shallow water equations \eqref{eq:sw}.

\subsection{Formation of the CHJ}\label{sec:formation}

\clearpage
\section{Numerical methods}
In this work, we consider Riemann-based finite volume methods.
Considering a one dimensional hyperbolic conservation law,
the cell average of the numerical solution given by a 
finite volume scheme (with forward Euler) can be written as follows:
\begin{align}\label{first-order_FV}
  \bfu_i^{n+1}=\bfu_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right],
\end{align}
where $\bfF_{i+1/2}=\bfF_{i+1/2}(u^n)$ is a numerical flux; i.e., $\bfF_{i+1/2}$ is 
a discretization of the flux $\bff(\bfu(x_{i+1/2},t^n))$. 
This numerical flux is composed of two parts, a consistent term that approximates 
the physical flux at the interface between cells $i$ and $i+1$, and a stabilization term. 
For instance, the well-known Rusanov's numerical flux is given by 
\begin{align}\label{rusanov_num_flux}
  \bfF_{i+1/2}(\bfu)=\frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2} - \lambda_{i+1/2}(\bfu_{i+1}-\bfu_i),
\end{align}
where $\lambda_{i+1/2}$ is an upper bound on the wave speed associated with the Riemann problem 
composed of the states $\bfu_i$ and $\bfu_{i+1}$. 
%If $z=u$ in \eqref{rusanov_num_flux}
%; i.e., if the Rusanov's numerical flux is obtained considering 
%the cell averages of the solution 
%and if the time step is small enough, 
Method \eqref{first-order_FV} with Rusanov's flux \eqref{rusanov_num_flux}
is positivity preserving and entropy stable, see for instance 
\cite{perthame1996positivity,tadmor2003entropy,guermond2016invariant}
However, the accuracy of the numerical flux, and hence of the method, is 
limited to first-order. 
Two alternatives are common to improve the accuracy properties of Rusanov's method.
The numerical flux $\bfF_{i+1/2}$ can be approximated based on high-order polynomial 
reconstructions (or combinations of them) around the interface $i+1/2$. 
For smooth solutions, this strategy leads to arbitrarily high-order
methods. Alternatively, second-order accuracy can be obtained by removing some of the 
numerical dissipation introduced by the stabilization component of the numerical flux. 
The second-order Lax-Wendroff method \cite{lax1959systems} is a popular example of this approach. 
In this work we consider an adaptation of Lax-Wendroff method,
sometimes known as Lax-Wendroff-LeVeque \cite{leveque1997wave,leveque2002finite}, 
that incorporates limiters into the second-order correction. 
These high-order corrections to Rusanov's method require careful treatment to guarantee 
the solution is still positivity preserving and/or entropy stable.

The accuracy of a given first- or second-order method depends heavily on the Riemann solver. 
In particular, the Rusanov's Riemann solver is highly dissipative. In contrast, the well-known 
Roe's Riemann solver produces more accurate solutions. 
%
As we discussed in the introduction, finite volume methods with 
Roe's Riemann solvers are prone to the carbuncle instability for the CHJ. 
In contrast, if we use the Rusanov's 
Riemann solver, no carbuncles appear; however, many other important and physical features of 
the solution are dissipated, unless the grid is highly refined. 
Our main objective is to study the CHJ and to propose a Riemann solver
that is carbuncle free but do not dissipate important features of the solution. 
To do that we propose a Riemann solver that blends Rusanov's and Roe's solvers. 
Our blending criteria is based on the residual of the entropy (in)equality,
which induces dissipation of entropy near discontinuities. 
Blending these Riemann solvers has been proposed before
to solve the Euler equations, see e.g.
\cite{nishikawa2008very,wang2016developing,jaisankar2007diffusion,ohwada2018simple,deng2019new,ray2013entropy},
the shallow water equations, see e.g. \cite{bader2014carbuncle,kemm2014note},
and even the Navier-Stokes equations, see e.g. \cite{nishikawa2008very,ohwada2018simple}.
In these references, the authors use blending criteria different than ours.
Nevertheless, fixing the carbuncle instability by imposing entropy stability is not new.
For example, in \cite{ismail2009affordable,ismail2009proposed},
the authors developed carbuncle free solvers for the Euler equations
via dissipation of entropy.

@@

The finite volume method \eqref{first-order_FV} can be written in terms of right- 
and left-going waves emanating from the interfaces $i-1/2$ and $i+1/2$, respectively. 
The net contribution of these waves is called right- and left-going fluctuations. 
Upon defining $z^+:=\frac{1}{2}(z+|z|)$ and $z^-:=\frac{1}{2}(z-|z|)$, 
the fluctuations are given by
\begin{align}\label{fluct}
  %\A^+\Delta Q_{i-1/2} := \frac{1}{2}\sum_p\left(s_{i-1/2}^p+|s_{i-1/2}^p|\right)\W_{i-1/2}^p, \quad
  %\A^-\Delta Q_{i+1/2} := \frac{1}{2}\sum_p\left(s_{i+1/2}^p-|s_{i+1/2}^p|\right)\W_{i+1/2}^p,
  \A^+\Delta Q_{i-1/2} := \sum_p\left(s_{i-1/2}^p\right)^+\W_{i-1/2}^p, \qquad
  \A^-\Delta Q_{i+1/2} := \sum_p\left(s_{i+1/2}^p\right)^-\W_{i+1/2}^p,
\end{align}
where $\W_{i\pm 1/2}^p$ and $s_{i\pm 1/2}^2$ are the waves and their corresponding speeds, respectively. 
The waves and their speed of propagation (to compute the fluctuations) are tightly related to Riemann solvers. 
In \S\ref{sec:riemann_solvers} we review the definition of the waves and speeds associated with 
Rusanov's and Roe's Riemann solvers.
Based on fluctuations, method \eqref{first-order_FV} can be written as 
\begin{align}\label{first-order_via_fluct}
  u_i^{n+1}=u_i^n-\frac{\Delta t}{\Delta x}\left[\A^+\Delta Q_{i-1/2}+\A^-\Delta Q_{i+1/2}\right].
\end{align}

The second-order Lax-Wendroff-LeVeque method removes some of the numerical dissipation in 
\eqref{first-order_via_fluct} via a flux correction at each interface. The method is written as follows:
\begin{align}\label{second-order_via_fluct}
  u_i^{n+1} = u_i^n-\frac{\Delta t}{\Delta x}\left[\A^+\Delta Q_{i-1/2}+\A^-\Delta Q_{i+1/2}\right]
  -\frac{\Delta t}{\Delta x}\left[\tilde{F}_{i+1/2}-\tilde{F}_{i-1/2}\right],
\end{align}
where 
\begin{align*}
  \tilde{F}_{i\pm 1/2}=\frac{1}{2}\sum_p|s_{i\pm 1/2}^p|\left(1-\frac{\Delta t}{\Delta x}|s_{i\pm 1/2}^p|\right)\tilde\W_{i\pm 1/2}^p
\end{align*}
are the flux corrections at the interfaces $i\pm 1/2$. Here $\tilde{\W}_{i\pm 1/2}^p$ is a 
TVD-limited version of $\W_{i\pm 1/2}^p$. 
In this work, we consider the min-mod limiters, we refer to \cite{leveque1997wave} for more details. 
If no limiters are used 
($\tilde \W_{i\pm 1/2}^p\equiv \W_{i\pm 1/2}^p$), \eqref{second-order_via_fluct} becomes
the standard second-order Lax-Wendroff scheme, which in terms of fluctuation-like quantities 
is given by
\begin{align*}
  u_i^{n+1}=u_i^n-\frac{\Delta t}{\Delta x}
\left[
  \frac{1}{2}\left(s_{i-1/2}^p+\frac{\Delta t}{\Delta x}|s_{i-1/2}^p|^2\right)\W_{i-1/2}^p
  +
  \frac{1}{2}\left(s_{i+1/2}^p-\frac{\Delta t}{\Delta x}|s_{i+1/2}^p|^2\right)\W_{i+1/2}^p
\right].
\end{align*}
To implement method \eqref{second-order_via_fluct}, 
we need to choose a Riemann solver. Although the scheme is second-order 
regardless of the Riemann solver, the accuracy and robustness depends upon the 
Riemann solver. The approximate Riemann solver based on Rusanov's flux is robust and known to be carbuncle free. 
However, it adds large amounts of numerical diffusion that, for relative coarse grids, 
might dissipate other physical and important features of the solution. If the apprroximate Riemann solver 
is based on the Roe's average, the amount of numerical diffusion is highly reduced; however, 
such an approach is prone to develop the carbuncle instability for certain problems. 
Many carbuncle fixes to the Roe's solver have been proposed for the Euler equations. 
In the context of the shallow water equations, we review in \S\ref{sec:riemann_solvers} 
the Riemann solver proposed in \cite{kemm2014note}.

@@

\section{Entropy dissipative Riemann solver}\label{sec:ent_diss_rs}
Consider a hyperbolic system of conservation laws 
\begin{align}\label{cons_law}
  \frac{\partial \bfu(\bfx)}{\partial t} + \nabla\cdot \bff(\bfu(\bfx))=0, \quad \bfx\in\Omega,
\end{align}
with a en entropy pair $\left(\eta(\bfu), \bfq(\bfu)\right)$. 
A weak solution of \eqref{cons_law} is called an entropy solution if 
\begin{align}\label{ent_inequality}
  \frac{\partial \eta(\bfu(\bfx))}{\partial t} + \nabla\cdot \bfq(\bfu(\bfx))\leq 0, \quad \bfx\in\Omega,
\end{align}
holds, in a distribution sense, for any entropy pair.
It is desirable for a numerical method to satisfy a discrete version of the entropy inequality 
\eqref{ent_inequality}. This is, however, not true for many standard numerical discretizations. 
In general, an entropy stable scheme introduces nonlinear stabilization, which is 
particularly important for solutions with strong discontinuities that might converge to non-physical 
weak solutions. 
In this work, we propose an entropy stable approximate Riemann solver for the one dimensional 
shallow water equations. We use this Riemann solver as building block within the Lax-Wendroff-LeVeque 
method, as described in \cite{leveque1997wave,leveque2002finite} and implemented in Clawpack.

@@

\subsection{Entropy stable Riemann solver}\label{sec:entropy_stable}
The Riemann solver presented in the previous section blends Rusanov's and Roe's solver based on a
measurement of the entropy residual. The objective is to induce dissipation of entropy 
near discontinuities. However, we have no guarantee that methods \eqref{first-order_via_fluct} 
or \eqref{second-order_via_fluct} based on this Riemann solver are (locally) entropy stable. 
In this section, we consider the one-dimensional problem and increase the numerical dissipation 
to guarantee entropy stability for \eqref{first-order_via_fluct}. 
To this end we follow the work by Tadmor, see for instance \cite{tadmor1987numerical, tadmor2003entropy} 
and references therein.

>>>>

\newcommand{\W}{{\mathcal W}}
\newcommand{\A}{{\mathcal A}}
\newcommand{\apdq}{\A^+ \!\!{\Delta} Q}
\newcommand{\amdq}{\A^- \!\!{\Delta} Q}
\newcommand{\imh}{{i-1/2}}
\newcommand{\iph}{{i+1/2}}
\newcommand{\bff}{{\bf f}}
\newcommand{\bfr}{{\bf r}}
\newcommand{\bfF}{{\bf F}}
\newcommand{\bfu}{{\bf u}}
\newcommand{\bfv}{{\bf v}}
\newcommand{\bfq}{{\bf q}}
\newcommand{\bfx}{{\bf x}}
\newcommand{\sgn}{{\rm sgn}}
\newcommand{\Rus}{{\rm Rus}}
\newcommand{\Roe}{{\rm Roe}}

@@

\section{The shallow water circular hydraulic jump}
We consider the shallow water model in two horizontal dimensions:
\begin{subequations} \label{eq:sw}
\begin{align}
    h_t + (hu)_x + (hv)_y & = 0 \\
    (hu)_t + \left(hu^2 + \frac{1}{2}gh^2\right)_x + (huv)_y & = 0 \\
    (hv)_t + (huv)_x + \left(hv^2 + \frac{1}{2}gh^2\right)_y & = 0.
\end{align}
\end{subequations}
Here $h, u,$ and $v$ are respectively the depth and the $x$- and $y$-components of
velocity, which are functions of the two spatial coordinates $(x,y)$ as well as time $t$.
The gravitational force is proportional to $g$.
%The system \eqref{eq:sw} can be written in vector form as
%\begin{align}
%    q_t + f(q)_x + g(q)_y & = 0.
%\end{align}

\subsection{Semi-analytical steady solution under rotational symmetry}\label{sec:steady_chj}
In this work, we consider the initial boundary value problem consisting of the
shallow water model \eqref{eq:sw} in an annular domain
\begin{align}
r_\textup{jet} \le r \le r_\infty
\end{align}
where $r = \sqrt{x^2+y^2}$, with prescribed inflow at $r=r_\textup{jet}$
and prescribed outflow at $r=r_\infty$.
The domain and boundary conditions are rotationally symmetric.
By assuming rotational symmetry in \eqref{eq:sw}, one obtains the system
\begin{subequations} \label{eq:rsw}
\begin{align}
    (rh)_t + (rhu)_r & = 0 \label{mass1} \\
    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r = 0. \label{mom1}
\end{align}
\end{subequations}
where the depth $h$ and radial velocity $u$ are functions of the single spatial coordinate $r$.
Steady-state solutions of \eqref{eq:rsw} satisfy
\begin{subequations}\label{steady}
\begin{align}
    rhu & = C \\
    h'(r) & = \frac{h}{\frac{g}{\beta^2} r^3 h^3 -r} = \frac{h}{r} \cdot \frac{F^2}{1-F^2} \label{eq:dh}
\end{align}
\end{subequations}
for some $C$ independent of $r$.  Here $F=|u|/\sqrt{gh}$ is the Froude number.
We see that two types of steady profiles exist, depending on whether the flow
is subcritical ($|F|<1$) or supercritical ($|F|>1$).  No smooth steady solution can
include both regimes, since the right hand side of \eqref{eq:dh} blows up when $F=1$.

@@

A steady-state, rotationally symmetric solution can be given for an annular region with prescribed
flow at the inner and outer boundaries as follows:

@@

Numerical tests suggest that the solution of \eqref{eq:rsw} tends rapidly approaches that
just described under general initial conditions as long as the flow at the jet is
supercritical and the outer outflow is subcritical.  This solution remains steady
if the rate of inflow and outflow are matched.  The stability of this solution in
the face of non-rotationally-symmetric perturbations is an important question not
only in the shallow water context but for more realistic fluid models and physically.
It will play an important role in the results we present later.
The rotationally-symmetric steady state is a useful initial condition for studies
of CHJ stability.

\section{Numerical methods}
In this work, we consider Riemann-based finite volume methods.
Considering a one dimensional hyperbolic conservation law,
the cell average of the numerical solution given by a 
finite volume scheme (with forward Euler) can be written as follows:
\begin{align}\label{first-order_FV}
  \bfu_i^{n+1}=\bfu_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right],
\end{align}
where $\bfF_{i+1/2}=\bfF_{i+1/2}(u^n)$ is a numerical flux; i.e., $\bfF_{i+1/2}$ is 
a discretization of the flux $\bff(\bfu(x_{i+1/2},t^n))$. 
This numerical flux is composed of two parts, a consistent term that approximates 
the physical flux at the interface between cells $i$ and $i+1$, and a stabilization term. 
For instance, the well-known Rusanov's numerical flux is given by 
\begin{align}\label{rusanov_num_flux}
  \bfF_{i+1/2}(\bfu)=\frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2} - \lambda_{i+1/2}(\bfu_{i+1}-\bfu_i),
\end{align}
where $\lambda_{i+1/2}$ is an upper bound on the wave speed associated with the Riemann problem 
composed of the states $\bfu_i$ and $\bfu_{i+1}$. 
%If $z=u$ in \eqref{rusanov_num_flux}
%; i.e., if the Rusanov's numerical flux is obtained considering 
%the cell averages of the solution 
%and if the time step is small enough, 
Method \eqref{first-order_FV} with Rusanov's flux \eqref{rusanov_num_flux}
is positivity preserving and entropy stable, see for instance 
\cite{perthame1996positivity,tadmor2003entropy,guermond2016invariant}
However, the accuracy of the numerical flux, and hence of the method, is 
limited to first order. 
Two alternatives are common to improve the accuracy properties of Rusanov's method.
The numerical flux $\bfF_{i+1/2}$ can be approximated based on high-order polynomial 
reconstructions (or combinations of them) around the interface $i+1/2$. 
For smooth solutions, this strategy leads to arbitrarily high-order
methods. Alternatively, second-order accuracy can be obtained by removing some of the 
numerical dissipation introduced by the stabilization component of the numerical flux. 
The second-order Lax-Wendroff method \cite{lax1959systems} is a popular example of this approach. 
In this work we consider an adaptation of Lax-Wendroff method,
sometimes known as Lax-Wendroff-LeVeque \cite{leveque1997wave,leveque2002finite}, 
that incorporates limiters into the second-order correction. 
These high-order corrections to Rusanov's method require careful treatment to guarantee 
the solution is still positivity preserving and/or entropy stable.

The accuracy of a given first- or second-order method depends heavily on the Riemann solver. 
In particular, the Rusanov's Riemann solver is highly dissipative. In contrast, the well-known 
Roe's Riemann solver produces more accurate solutions. 
%
As we discussed in the introduction, finite volume methods with 
Roe's Riemann solvers are prone to the carbuncle instability for the CHJ. 
In contrast, if we use the Rusanov's 
Riemann solver, no carbuncles appear; however, many other important and physical features of 
the solution are dissipated, unless the grid is highly refined. 
Our main objective is to study the CHJ and to propose a Riemann solver
that is carbuncle free but do not dissipate important features of the solution. 
To do that we propose a Riemann solver that blends Rusanov's and Roe's solvers. 
Our blending criterion is based on the residual of the entropy (in)equality,
which induces dissipation of entropy near discontinuities. 
Blending these Riemann solvers has been proposed before
to solve the Euler equations, see e.g.
\cite{nishikawa2008very,wang2016developing,jaisankar2007diffusion,ohwada2018simple,deng2019new,ray2013entropy},
the shallow water equations, see e.g. \cite{bader2014carbuncle,kemm2014note},
and even the Navier-Stokes equations, see e.g. \cite{nishikawa2008very,ohwada2018simple}.
In these references, the authors use blending criteria different than ours.
Nevertheless, fixing the carbuncle instability by imposing entropy stability is not new.
For example, in \cite{ismail2009affordable,ismail2009proposed},
the authors developed carbuncle free solvers for the Euler equations
via dissipation of entropy.

@@

The finite volume method \eqref{first-order_FV} can be written in terms of right- 
and left-going waves emanating from the interfaces $i-1/2$ and $i+1/2$, respectively. 
The net contribution of these waves is called right- and left-going fluctuations. 
Upon defining $z^+:=\frac{1}{2}(z+|z|)$ and $z^-:=\frac{1}{2}(z-|z|)$, 
the fluctuations are given by
\begin{align}\label{fluct}
  %\A^+\Delta Q_{i-1/2} := \frac{1}{2}\sum_p\left(s_{i-1/2}^p+|s_{i-1/2}^p|\right)\W_{i-1/2}^p, \quad
  %\A^-\Delta Q_{i+1/2} := \frac{1}{2}\sum_p\left(s_{i+1/2}^p-|s_{i+1/2}^p|\right)\W_{i+1/2}^p,
  \apdq_\imh := \sum_p\left(s_{i-1/2}^p\right)^+\W_{i-1/2}^p, \qquad
  \amdq_\iph := \sum_p\left(s_{i+1/2}^p\right)^-\W_{i+1/2}^p, \\
\end{align}
where $\W_{i\pm 1/2}^p$ and $s_{i\pm 1/2}^2$ are the waves and their corresponding speeds, respectively. 
The waves and their speed of propagation (to compute the fluctuations) are tightly related to Riemann solvers. 
In \S\ref{sec:riemann_solvers} we review the definition of the waves and speeds associated with 
Rusanov's and Roe's Riemann solvers.
Based on fluctuations, method \eqref{first-order_FV} can be written as 
\begin{align}\label{first-order_via_fluct}
  u_i^{n+1}=u_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right].
\end{align}

The second-order Lax-Wendroff-LeVeque method removes some of the numerical dissipation in 
\eqref{first-order_via_fluct} via a flux correction at each interface. The method is written as follows:
\begin{align}\label{second-order_via_fluct}
  u_i^{n+1} = u_i^n-\frac{\Delta t}{\Delta x}\left[\A^+\Delta Q_{i-1/2}+\A^-\Delta Q_{i+1/2}\right]
  -\frac{\Delta t}{\Delta x}\left[\tilde{F}_{i+1/2}-\tilde{F}_{i-1/2}\right],
\end{align}
where 
\begin{align*}
  \tilde{F}_{i\pm 1/2}=\frac{1}{2}\sum_p|s_{i\pm 1/2}^p|\left(1-\frac{\Delta t}{\Delta x}|s_{i\pm 1/2}^p|\right)\tilde\W_{i\pm 1/2}^p
\end{align*}
are the flux corrections at the interfaces $i\pm 1/2$. Here $\tilde{\W}_{i\pm 1/2}^p$ is a 
TVD-limited version of $\W_{i\pm 1/2}^p$. 
In this work, we consider the min-mod limiters, we refer to \cite{leveque1997wave} for more details. 
If no limiters are used 
($\tilde \W_{i\pm 1/2}^p = \W_{i\pm 1/2}^p$), \eqref{second-order_via_fluct} becomes
the standard second-order Lax-Wendroff scheme, which in terms of fluctuation-like quantities 
is given by
\begin{align*}
  u_i^{n+1}=u_i^n-\frac{\Delta t}{\Delta x}
\left[
  \frac{1}{2}\left(s_{i-1/2}^p+\frac{\Delta t}{\Delta x}|s_{i-1/2}^p|^2\right)\W_{i-1/2}^p
  +
  \frac{1}{2}\left(s_{i+1/2}^p-\frac{\Delta t}{\Delta x}|s_{i+1/2}^p|^2\right)\W_{i+1/2}^p
\right].
\end{align*}
To implement method \eqref{second-order_via_fluct}, 
we need to choose a Riemann solver. Although the scheme is second-order 
regardless of the Riemann solver, the accuracy and robustness depends upon the 
Riemann solver. The approximate Riemann solver based on Rusanov's flux is robust and known to be carbuncle free. 
However, it adds large amounts of numerical diffusion that, for relatively coarse grids, 
might dissipate other physical and important features of the solution. If the approximate Riemann solver 
is based on the Roe's average, the amount of numerical diffusion is highly reduced; however, 
such an approach is prone to develop the carbuncle instability for certain problems. 
Many carbuncle fixes to the Roe's solver have been proposed for the Euler equations. 
In the context of the shallow water equations, we review in \S\ref{sec:riemann_solvers} 
the Riemann solver proposed in \cite{kemm2014note}.

@@

\section{Entropy-dissipative Riemann solver}\label{sec:ent_diss_rs}
Consider a hyperbolic system of conservation laws 
\begin{align}\label{cons_law}
  \frac{\partial \bfu(\bfx)}{\partial t} + \nabla\cdot \bff(\bfu(\bfx))=0, \quad \bfx\in\Omega,
\end{align}
with a en entropy pair $\left(\eta(\bfu), \bfq(\bfu)\right)$. 
A weak solution of \eqref{cons_law} is called an entropy solution if 
\begin{align}\label{ent_inequality}
  \frac{\partial \eta(\bfu(\bfx))}{\partial t} + \nabla\cdot \bfq(\bfu(\bfx))\leq 0, \quad \bfx\in\Omega,
\end{align}
holds, in a distribution sense, for any entropy pair.
It is desirable for a numerical method to satisfy a discrete version of the entropy inequality 
\eqref{ent_inequality}, but many widely-used discretizations do not.
In general, an entropy stable scheme introduces nonlinear stabilization, which is 
particularly important for solutions with strong discontinuities that might converge to non-physical 
weak solutions. 
In this work, we propose an entropy stable approximate Riemann solver for the one dimensional 
shallow water equations. We use this Riemann solver as building block within the Lax-Wendroff-LeVeque 
method, as described in \cite{leveque1997wave,leveque2002finite} and implemented in Clawpack.

@@

\subsection{Entropy-stable Riemann solver}\label{sec:entropy_stable}
The Riemann solver presented in the previous section blends Rusanov's and Roe's solver based on a
measurement of the entropy residual. The objective is to induce dissipation of entropy 
near discontinuities. However, we have no guarantee that methods \eqref{first-order_via_fluct} 
or \eqref{second-order_via_fluct} based on this Riemann solver are (locally) entropy stable. 
In this section, we consider the one-dimensional problem and increase the numerical dissipation 
to guarantee entropy stability for \eqref{first-order_via_fluct}. 
To this end we follow the work by Tadmor, see for instance \cite{tadmor1987numerical, tadmor2003entropy} 
and references therein.
